Point-process based representation learning for Electronic Health Records

### 2023-01-26 Manuscript_v1.tex
commit
com

\textbf{Synthea (\emph{Syn})}.
We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes. Here, we reused the already processed version of this data by Edgauard.

@@

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage{graphicx}
\begin{table}[h!]
    \centering
    \caption{unsupervised learning}
    \label{tab:table1}
    \resizebox{\columnwidth}{!}{%    
    \begin{tabular}{cclllll}
    \hline
    \multirow{2}{*}{Dataset} & \multicolumn{2}{c}{Model}            & \multicolumn{2}{c}{next mark} & next time & CIFs  \\
                             & \multicolumn{1}{l}{module} & loss    & AUPRC         & AUROC         & RMSE      & NLL   \\ \hline
    \multirow{6}{*}{p12}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
    \multirow{6}{*}{p19}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
    \end{tabular}%
    }
\end{table}

@@

\subsection{Next event prediction}

using CIF does not seem to improve the next mark and time prediction.
this is new insight as previous studies has claimed this loss to be useful.

\subsection{Learned representation}

using CIF loss function leads to a better representation for sepsis.

@@

Although CIF does not imporve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.

In addition, we can interprete some of learned CIF patterns.

@@

one advantage of proposed method is use of attention mechanisms in both event and state encoder.

>>>>

\textbf{Synthea(\emph{Syn})}.
We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes. Here, we reused the already processed version of this data by Edgauard.

@@

In this section, we present our results regarding the advantage of state and event encoding.

\subsection{Negative Likelihood with state encoding}

Table 1 shows the result for estimation of negative likelihood in different datasets and scnearios. It is obvious that state encoding has led to lower NLL.

\subsection{Time concatenation}

\subsection{Downstream task with event encoding}

Another key element of our work is to show the effectiveness of point process modeling for a down-stream task. In Table, we have reported the performance metrics for the mortality prediction task. We have compared our results with several sota's DL models that are compatible with irregular time series.

\subsection{Learned representions}
Fig 1 visualizes the tsne plot for the two scenarios.

one advantage of proposed method is use of attention mechanisms in both event and state encoder.
Fig 1 shows the attention mechanism
% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage{graphicx}
\begin{table}[h!]
    \centering
    \caption{unsupervised learning}
    \label{tab:table1}
    \resizebox{\columnwidth}{!}{%    
    \begin{tabular}{cclllll}
    \hline
    \multirow{2}{*}{Dataset} & \multicolumn{2}{c}{Model}            & \multicolumn{2}{c}{next mark} & next time & CIFs  \\
                             & \multicolumn{1}{l}{module} & loss    & AUPRC         & AUROC         & RMSE      & NLL   \\ \hline
    \multirow{6}{*}{p12}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
    \multirow{6}{*}{p19}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
    \end{tabular}%
    }
\end{table}

@@

@@

Although CIF does not improve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.

In addition, we can interpret some of learned CIF patterns.

@@
